---
title: "clinvaR vignette"
author: "James Diao"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = F}
knitr::opts_knit$set(root.dir = ".");
knitr::opts_chunk$set(echo = T, eval = T, cache = F, warning = F, message = F)
```

This is a vignette intended to introduce new users to `clinvaR` ([https://github.com/jamesdiao/clinvaR](https://github.com/jamesdiao/clinvaR)). 

## Purpose
For `clinvaR`, the first goal is to go from genes to a vcf of related variants, along with classifications from any version of ClinVar.  
This involves 3 steps:  

1. Select a list of genes that you are interested in. clinvaR makes it easy to download and import relevant variants from 1KG.  
2. Select a version of ClinVar to assign pathogenicity labels with. All previous ClinVar VCFs are stored as binary files and can be retrieved with a "closest date" function.  
3. Select an analysis method. This part is not ready yet.  

## Dependencies

clinvaR has a number of file dependencies that are generally useful for these types of projects. I have attached them here. 
```{r, echo = F}
pkg_list <- c("remotes","pander","ggplot2","tibble","tidyr","dplyr", "stringr")
installed <- installed.packages()[,"Package"]
is.installed <- pkg_list %in% installed
if (!all(is.installed))
  install.packages(pkg_list[!is.installed])
print(names(sapply(pkg_list, require, character.only = T)), quote = F)
if (!("clinvaR" %in% installed))
  remotes::install_github('jamesdiao/clinvaR')
require(clinvaR)
```


## Select a List of Genes

The first step is to have a list of genes that you are interested in. 

### Method 1: Direct assignment
The easiest way is to directly save your genes of interest in gene_panel as a character vector (all caps). We have taken the example of the 20 genes on the Laboratory of Molecular Medicine's gene testing panel for hypertrophic cardiomyopathy. 


```{r}
gene_panel <- c("ACTC1", "ACTN2", "CSRP3", "GLA", "LAMP2", "MYBPC3", "MYH7", 
"MYL2", "MYL3", "MYOZ2", "NEXN", "PLN", "PRKAG2", "PTPN11", "RAF1", 
"TNNC1", "TNNI3", "TNNT2", "TPM1", "TTR")
```

### Method 2: Import from text file

#### From own file
Another way is to save your gene list in a text file (1 gene per line) and import it. 

```{r}
path <- system.file("extdata/MacArthur_Gene_Lists/lists/lmm_hcm.tsv", package = "clinvaR")
print(path)
```

```{r}
gene_panel <- get_genes(path)
print(gene_panel)
```

#### From clinvaR stored lists
The MacArthur Lab has assembled a compilation of gene lists ([https://github.com/macarthur-lab/gene_lists](https://github.com/macarthur-lab/gene_lists)) that may be a useful starting point. We have included all sets (and a few of our own) with abbreviated descriptions. Please visit the GitHub repo for citation information. 

| List | File Name | Count | Description |
| ---------------- | ------- | ------- | ------------------------------------------------------- | 
| Universe | universe.tsv | 18,991 | Approved symbols for 18,991 protein-coding genes according to HGNC. All other lists are subsets. |
| Hypertrophic cardiomyopathy | lmm_hcm.tsv | 20 | Gene testing panel for hypertrophic cardiomyopathy from the Laboratory of Molecular Medicine |
| ACMG secondary findings recommendations v2.0 | acmg_59.tsv | 59 | The American College of Medical Genetics and Genomics has recommended that sequencing laboratories seek and report secondary findings in a minimum list of 59 genes. |
| BROCA - Cancer Risk Panel | BROCA_Cancer_Risk_Panel.tsv | 66 | Suspected hereditary cancer predisposition, with a focus on breast or ovarian cancers. May co-occur with other cancer types (such as colorectal, endometrial, pancreatic, endocrine, or melanoma). |
| DNA Repair Genes, KangJ| DRG_KangJ.tsv | 151 | 151 DNA repair genes from DNA repair pathways: ATM, BER, FA/HR, MMR, NHEJ, NER, TLS, XLR, RECQ, and other. | 
| ClinGen haploinsufficient genes | clingen_level3_genes_2015_02_27.tsv | 221 | Genes with sufficient evidence for dosage pathogenicity (level 3) as determined by the ClinGen Dosage Sensitivity Map |
| Genes with any disease association reported in ClinVar | clinvar_path_likelypath.tsv | 3078 | All gene symbols for which there is at least one variant with an assertion of pathogenic or likely pathogenic in ClinVar. | 
| FDA-approved drug targets | fda_approved_drug_targets.tsv | 286 | Genes whose protein products are known to be the mechanistic targets of FDA-approved drugs. |
| Drug targets by Nelson et al 2012 | drug_targets_nelson.tsv | 201 | Drug targets according to Nelson et al 2012. |
| X-linked ClinVar genes | x-linked_clinvar.tsv | 61 | X chromosome genes in the August 6, 2015 ClinVar release that have at least 3 reportedly pathogenic, non-conflicted variants in ClinVar with at least one submitter other than OMIM or GeneReviews.|
| All dominant genes | all_ad.tsv | 709 | Currently the union of the Berg and Blekhman autosomal dominant OMIM gene lists, may add more lists later. |
| All recessive genes | all_ar.tsv | 1183 | Currently the union of the Berg and Blekhman autosomal recessive OMIM gene lists, may add more lists later. | 
| Essential in culture | core_essentials_hart.tsv | 285 | Genes deemed essential in multiple cultured cell lines based on shRNA screen data |
| Essential in mice | mgi_essential.tsv | 2,454 | Genes where homozygous knockout in mice results in pre-, peri- or post-natal lethality. |
| Genes nearest to GWAS peaks | gwascatalog.tsv | 3,762 | Closest gene 3' and 5' of GWAS hits in the NHGRI GWAS catalog as of Feb 9, 2015 |
| Kinases | kinases.tsv | 351 | From UniProt's [pkinfam list](http://www.uniprot.org/docs/pkinfam) |
| G-protein-coupled receptors | gpcr.tsv | 1705 | GPCR list from [guidetopharmacology.org](http://www.guidetopharmacology.org/GRAC/GPCRListForward?class=A) | 
| Natural product targets| natural_product_targets.tsv | 37 | List of hand-curated targets of natural products |


These are easily imported by directly referring to the file name. 

```{r}
gene_panel <- get_genes('lmm_hcm.tsv')
print(gene_panel)
```

The last way is to search OMIM for related genes. 
```{r}
#omim_table <- read.table('clinvaR/storage/gene_lists/other_data/omim.full.tsv', 
#                         sep = '\t', fill = T, header = T)
#search_terms <- c("Cardiomyopathy")
```

Then, you can download variants of interest from 1000 Genomes

```{r}
#download_output <- download_1000g(genes = gene_panel)
#final_vcf <- import_file_1000g(genes = gene_panel)
```



## Import ClinVar VCF

This must be merged with some version of ClinVar. This can be done either using the latest saved version (July 5, 2017), or by specifying a date from the list. 

```{r}
get_date_list()
```

Use `get_clinvar()` to collect from a certain date (if date is not present, the closest date is used). 

```{r}
clinvar <- get_clinvar("2017-07-21")
clinvar[1:4,] #%>% select(-VAR_ID, -CLNDBN)
```


```{r}
clinvar <- get_clinvar()
clinvar[1:4,] #%>% select(-VAR_ID, -CLNDBN)
```


Then, we merge clinvar with the genes of interest. 

```{r, cache = T}
#merged_1000g <- merge_clinvar_1000g(clinvar = get_clinvar(Sys.Date()), 
#                    vcf = import_file_1000g(gene_panel))
#merged_1000g[1:5,1:5]
```

We can also slice clinvar itself to collect only relevant entries. 
```{r}
#clinvar$CHROM %>% paste0(str_pad(clinvar$POS, 10, side = 'left', pad = "0")) %>% as.numeric %>% 
#  between(start, stop) -> filter_condition
#clinvar <- filter(clinvar, filter_condition)
```


